---
title: "Activity Analysis"
author: "Chris Ward"
date: "Thursday, May 14, 2015"
output: html_document
---

This is walk through of my analysis of the activity data set
============================================================

This code reads in the data
```{r}
setwd("~/Training/Data")
dat <- read.table("Activity.csv", header = T, sep = ",", as.is= TRUE)
```

This code processes the data. It formats the interval as a factor, the date as a date. It also creates a column which computes which day of week each observation takes place. Separate data frames are made for weekdays and weekends which will be utilized later
```{r}
dat$interval <- as.factor(dat$interval)
dat$date <- as.Date(dat$date, format="%m/%d/%Y")
dat$day <- weekdays(dat$date) #Adds day of week
dat$weekday <- ifelse(weekdays(dat$date)=="Saturday" | weekdays(dat$date)=="Sunday", "weekend", "weekday") #Adds weekend/weekday column
dat_wday <- subset(dat, dat$weekday == "weekday") # creates a weekday subset dataframe
dat_wend <- subset(dat, dat$weekday == "weekend") # creates a weekend subset dataframe
```


This code will compute the toal steps per day and create a histogram
```{r}
steps_per_day <- list() #creates empty list
for (i in 1:length(unique(dat$date))){ #counts total steps per day
  steps_per_day <- cbind(steps_per_day, sum(dat$steps[dat$date==(unique(dat$date))[i]]))
}
steps_per_day <- sapply(steps_per_day, as.numeric)  #converts to numeric
# 2.
hist(steps_per_day, breaks= length(unique(dat$date))) #creates a histogram
#3. 
mean(steps_per_day, na.rm=T)
median(steps_per_day, na.rm=T)
```

This code analyzes the daily activity pattern, creates a time series plot and outputs the interval with the most average steps
```{r}
steps_per_int <- list() #creates empty list
for (i in 1:length(unique(dat$interval))){ #counts total steps per interval
  steps_per_int <- cbind(steps_per_int,sum(dat$steps[dat$interval==as.character((unique(dat$interval))[i])],na.rm=T))
}
steps_per_int <- sapply(steps_per_int, as.numeric)  #converts to numeric
steps_per_int <- steps_per_int/length(unique(dat$date))
plot(steps_per_int, unique(dat$intervals), type="l", xlab= "", ylab="") # creates a plot of steps per interval
mtext("Steps", side=2, line=2.75, cex=1.2, lwd=4)
mtext("Time Interval", side=1, line=3, cex=1.2)
max_steps<- max(steps_per_int, na.rm=T) #Caluculates the max steps
match(max_steps,steps_per_int)
```

This code returns the number of NA's and imputs values for them
For weekend days I have imputed the average weekend interval # of steps and for week days I have impute the average weekday interval # of steps.

```{r}
#1
summary(dat)[7,1] #returns the number of intervals that have NA's
#2
dat_imputed <- dat # creates copy of dat to imput values on
avg_step_wday_int <- mean(dat_wday$steps, na.rm=T) # average weekday int
avg_step_wend_int <- mean(dat_wend$steps, na.rm=T) # average weekend int
# 3
if(dat_imputed$weekday=="weekday"){
  dat_imputed$steps[is.na(dat_imputed$steps)] <-avg_step_wday_int
  } else {
  dat_imputed$steps[is.na(dat_imputed$steps)] <-avg_step_wend_int
}
```
I have hidden the results for this portion because it returns every value that has been imputed
```{r, results="hide"}
ifelse(dat_imputed$steps=="Saturday" | dat_imputed$day=="Sunday",
       dat_imputed$steps[is.na(dat_imputed$steps)] <- avg_step_wday_int,
       dat_imputed$steps[is.na(dat_imputed$steps)] <- avg_step_wend_int)
```

```{r}
wday_steps_per_int <- list() #creates empty list
for (i in 1:length(unique(dat_wday$interval))){ #counts avg steps per interval
  wday_steps_per_int <- cbind(wday_steps_per_int,sum(dat_wday$steps[dat_wday$interval==as.character((unique(dat_wday$interval))[i])],na.rm=T))
}
wday_steps_per_int <- sapply(wday_steps_per_int, as.numeric)  #converts to numeric
wday_steps_per_int <- wday_steps_per_int/length(unique(dat_wday$date))
wday_avgs <- data.frame(unique(dat_wday$interval), wday_steps_per_int, "weekday")

wend_steps_per_int <- list() #creates empty list
for (i in 1:length(unique(dat_wend$interval))){ #counts avg steps per interval
  wend_steps_per_int <- cbind(wend_steps_per_int,sum(dat_wend$steps[dat_wend$interval==as.character((unique(dat_wend$interval))[i])],na.rm=T))
}
wend_steps_per_int <- sapply(wend_steps_per_int, as.numeric)  #converts to numeric
wend_steps_per_int <- wend_steps_per_int/length(unique(dat_wend$date))
wend_avgs <- data.frame(unique(dat_wend$interval), wend_steps_per_int, "weekend")

```